### 2023-01-30 latex/bibliography.tex
no "available at"

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
%\begin{CSLReferences}{1}{0}
\begin{thebibliography}{9}

\bibitem[5stardata.info]{ref-11dInoyVF}
5-star Open Data. \href{http://5stardata.info/en/}{\emph{http://5stardata.info/en/}} (accessed 24 January 2023).

\end{thebibliography}

\leavevmode\vadjust pre{\hypertarget{ref-CJuBupCW}{}}%
. \href{https://blog.cloudflare.com/http-3-vs-http-2/}{\emph{https://blog.cloudflare.com/http-3-vs-http-2/}}

\leavevmode\vadjust pre{\hypertarget{ref-18P9txgeB}{}}%
. \href{https://www.rd-alliance.org/sites/default/files/Cordra.2022.pdf}{\emph{https://www.rd-alliance.org/sites/default/files/Cordra.2022.pdf}}

\leavevmode\vadjust pre{\hypertarget{ref-1BlxdxOwa}{}}%
14:00-17:00.ISO 16684-1:2019. \href{https://www.iso.org/standard/75163.html}{\emph{https://www.iso.org/standard/75163.html}} (accessed 24 January 2023b).

\leavevmode\vadjust pre{\hypertarget{ref-ikO7e0mh}{}}%
14:00-17:00.ISO/IEC 23009-1:2019. \href{https://www.iso.org/cms/render/live/en/sites/isoorg/contents/data/standard/07/93/79329.html}{\emph{https://www.iso.org/cms/render/live/en/sites/isoorg/contents/data/standard/07/93/79329.html}} (accessed 26 May 2022a).

\leavevmode\vadjust pre{\hypertarget{ref-11dInoyVF}{}}%
5-star Open Data. \href{http://5stardata.info/en/}{\emph{http://5stardata.info/en/}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-18LsXvKGJ}{}}%
Allcock W, Bresnahan J, Kettimuthu R, Link M. The Globus Striped GridFTP Framework and Server. In: \emph{ACM/IEEE SC 2005 Conference (SC'05)}. IEEE,. DOI: \href{https://doi.org/10.1109/sc.2005.72}{10.1109/sc.2005.72}.

\leavevmode\vadjust pre{\hypertarget{ref-11NpETLq1}{}}%
Anders I, Blanchi Ch, Broeder D, Hellström M, Islam Sh, Jejkal Th, Lannom L, Peters K, Quick R, Schlemmer A, Schwardmann U, Soiland-Reyes S, Strawn G, Uytvanck D van, Weiland C, Wittenburg P, Zwölf C. 2022a. \emph{\href{https://docs.google.com/document/d/1GAj-1owAAPDF7hVis2dYPPCOHiPHrIXIo4j-S5AdEFI/edit}{FAIR digital object overview and specifications}}. (internal draft): FDO Forum.

\leavevmode\vadjust pre{\hypertarget{ref-ljVViWCl}{}}%
Anders I, Hellström M, Islam S, Jejkal T, Lannom L, Schwardmann U, Wittenburg P. 2022b. \emph{\href{https://docs.google.com/document/d/1c2mZziq5pIPmLxMHLcYqlWrjYsc2ezGMXvp0E46iljo/edit}{FDO PID profiles \& attributes}}. (internal draft): FDO Forum.

\leavevmode\vadjust pre{\hypertarget{ref-UzQhqk0M}{}}%
Bahim C, Casorrán-Amilburu C, Dekkers M, Herczog E, Loozen N, Repanas K, Russell K, Stall S. 2020. The FAIR data maturity model: An approach to harmonise FAIR assessments. \emph{Data Science Journal} 19. DOI: \href{https://doi.org/10.5334/dsj-2020-041}{10.5334/dsj-2020-041}.

\leavevmode\vadjust pre{\hypertarget{ref-KbbW0kGT}{}}%
Belshe M, Peon R. 2015. \emph{Hypertext Transfer Protocol Version 2 (HTTP/2)}. RFC Editor. DOI: \href{https://doi.org/10.17487/rfc7540}{10.17487/rfc7540}.

\leavevmode\vadjust pre{\hypertarget{ref-RRtwPF7Y}{}}%
Berners-Lee T, Fielding R, Masinter L. 2005. \emph{Uniform Resource Identifier (URI): Generic Syntax}. RFC Editor. DOI: \href{https://doi.org/10.17487/rfc3986}{10.17487/rfc3986}.

\leavevmode\vadjust pre{\hypertarget{ref-Guy7cjVY}{}}%
Berners-Lee T, Fischetti M. 1999. \emph{Weaving the Web: the original design and ultimate destiny of the World Wide Web by its inventor}. San Francisco: HarperSanFrancisco.

\leavevmode\vadjust pre{\hypertarget{ref-117tX9MEG}{}}%
Bernstein A, Hendler J, Noy N. 2016. A new look at the semantic web. \emph{Communications of the ACM} 59:35--37. DOI: \href{https://doi.org/10.1145/2890489}{10.1145/2890489}.

\leavevmode\vadjust pre{\hypertarget{ref-1F8S9FeHf}{}}%
Bioschemas - Bioschemas. \href{http://bioschemas.org/}{\emph{http://bioschemas.org/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-19s3yyxDn}{}}%
Bizer C, Heath T, Berners-Lee T. 2009. Linked data - the story so far. \emph{International journal on Semantic Web and information systems} 5:1--22. DOI: \href{https://doi.org/10.4018/jswis.2009081901}{10.4018/jswis.2009081901}.

\leavevmode\vadjust pre{\hypertarget{ref-1H7cmmr69}{}}%
Blanchi C, Broeder D, Jejkal T, Sharif I, Schlemmer A, Uytvanck D van, Wittenburg P. 2022a. \emph{\href{https://docs.google.com/document/d/1fDR5VHbVIa2AbLsBR58idrfn_Ib3x6Fk-_LJ4c_Ftt4/edit}{FDO -- upload of FDO}}. (internal draft): FDO Forum.

\leavevmode\vadjust pre{\hypertarget{ref-126uxr5pI}{}}%
Blanchi C, Hellström M, Lannom L, Schwardmann U, Wittenburg P. 2022b. \emph{\href{https://docs.google.com/document/d/1RrOiwMhkl-hRzWmlluA2iXCzHK-bj7_80LlMXgWx4w4/edit}{Implementation of attributes, types, profiles and registries}}. (internal draft): FDO Forum.

\leavevmode\vadjust pre{\hypertarget{ref-EJDjHrUv}{}}%
Bonino Da Silva Santos LO, Wilkinson MD, Kuzniar A, Kaliyaperumal R, Thompson M, Dumontier M, Burger K. 2016. \href{https://www.researchgate.net/publication/\%7B309468587/_FAIR/_Data/_Points/_Supporting/_Big/_Data/_Interoperability\%7D}{FAIR data points supporting big data interoperability}. In: Zelm M, Doumeingts G, Mendonça JP eds. \emph{Enterprise interoperability in the digitized and networked factory of the future}. iSTE Press, 270--279.

\leavevmode\vadjust pre{\hypertarget{ref-RwvirqWg}{}}%
Bonino L, Wittenburg O, Carroll B, Hardisty A, Leggott M, Zwölf C. 2019. \emph{\href{https://github.com/GEDE-RDA-Europe/GEDE/blob/master/FAIR\%20Digital\%20Objects/FDOF/FAIR\%20Digital\%20Object\%20Framework-v1-02.docx}{FAIR digital object framework}}.

\leavevmode\vadjust pre{\hypertarget{ref-TUmA4KQT}{}}%
Carriero VA, Daquino M, Gangemi A, Nuzzolese AG, Peroni S, Presutti V, Tomasi F. 2020. The landscape of ontology reuse approaches. In: Cota G, Daquino M, Pozzato GL eds. \emph{Applications and practices in ontology design, extraction, and reasoning}. Studies on the semantic web. IOS Press,. DOI: \href{https://doi.org/10.3233/SSW200033}{10.3233/SSW200033}.

\leavevmode\vadjust pre{\hypertarget{ref-1E4OfHyhI}{}}%
Ciccarese P, Soiland-Reyes S, Belhajjame K, Gray AJ, Goble C, Clark T. 2013. PAV ontology: provenance, authoring and versioning. \emph{Journal of Biomedical Semantics} 4:37. DOI: \href{https://doi.org/10.1186/2041-1480-4-37}{10.1186/2041-1480-4-37}.

\leavevmode\vadjust pre{\hypertarget{ref-hRUsaTiz}{}}%
Clemm G, Amsden J, Ellison T, Kaler C, Whitehead J. 2002. \emph{Versioning Extensions to WebDAV (Web Distributed Authoring and Versioning)}. RFC Editor. DOI: \href{https://doi.org/10.17487/rfc3253}{10.17487/rfc3253}.

\leavevmode\vadjust pre{\hypertarget{ref-n2f3TfoI}{}}%
Content negotiation - HTTP \textbar{} MDN. \href{https://developer.mozilla.org/en-US/docs/Web/HTTP/Content_negotiation}{\emph{https://developer.mozilla.org/en-US/docs/Web/HTTP/Content\_negotiation}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-aCye3KpE}{}}%
Corcho O, Eriksson M, Kurowski K, Ojsteršek M, Choirat C, Sanden M van de, Coppens F. 2021. \emph{EOSC interoperability framework}. Publications Office of the EU. DOI: \href{https://doi.org/10.2777/620649}{10.2777/620649}.

\leavevmode\vadjust pre{\hypertarget{ref-o4h92LpS}{}}%
Data - W3C. \href{https://www.w3.org/standards/semanticweb/data}{\emph{https://www.w3.org/standards/semanticweb/data}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-zt8AAjcZ}{}}%
Data Catalog Vocabulary (DCAT) - Version 2. \href{https://www.w3.org/TR/2020/REC-vocab-dcat-2-20200204/}{\emph{https://www.w3.org/TR/2020/REC-vocab-dcat-2-20200204/}} (accessed 24 January 2023b).

\leavevmode\vadjust pre{\hypertarget{ref-lnH0hEuh}{}}%
Data Catalog Vocabulary (DCAT) - Version 2. \href{https://www.w3.org/TR/vocab-dcat-2/}{\emph{https://www.w3.org/TR/vocab-dcat-2/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-1EW4hLN8U}{}}%
Data Catalog Vocabulary (DCAT) - Version 3. \href{https://www.w3.org/TR/2022/WD-vocab-dcat-3-20220510/}{\emph{https://www.w3.org/TR/2022/WD-vocab-dcat-3-20220510/}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-6LvydTsC}{}}%
Data Information View. \href{https://handle-esgf.dkrz.de/lp/21.14100/2fcf49d3-0608-3373-a47f-0e721b7eaa87}{\emph{https://handle-esgf.dkrz.de/lp/21.14100/2fcf49d3-0608-3373-a47f-0e721b7eaa87}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-hKblceo6}{}}%
DCMI Metadata Terms. 2020. \href{https://www.dublincore.org/specifications/dublin-core/dcmi-terms/2020-01-20/}{\emph{https://www.dublincore.org/specifications/dublin-core/dcmi-terms/2020-01-20/}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-wAbZpOsr}{}}%
DCMI Metadata Terms. \href{https://www.dublincore.org/specifications/dublin-core/dcmi-terms/}{\emph{https://www.dublincore.org/specifications/dublin-core/dcmi-terms/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-KlCFFFyL}{}}%
Delgado JCM. 2016. An Interoperability Framework and Distributed Platform for Fast Data Applications. In: \emph{Data Science and Big Data Computing}. Springer International Publishing, 3--39. DOI: \href{https://doi.org/10.1007/978-3-319-31861-5_1}{10.1007/978-3-319-31861-5\_1}.

\leavevmode\vadjust pre{\hypertarget{ref-BFMoe8FC}{}}%
Designing a Linked Data developer experience. 2018. \href{https://ruben.verborgh.org/blog/2018/12/28/designing-a-linked-data-developer-experience/}{\emph{https://ruben.verborgh.org/blog/2018/12/28/designing-a-linked-data-developer-experience/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-16uB3jxpa}{}}%
Digital Object Interface Protocol Version 1.0 \textbar{} DONA Foundation. \href{https://www.dona.net/doipv1doc}{\emph{https://www.dona.net/doipv1doc}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-68fzID2k}{}}%
DOI Handbook - Resolution. \href{https://www.doi.org/doi_handbook/3_Resolution.html}{\emph{https://www.doi.org/doi\_handbook/3\_Resolution.html}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-1H9iQhQYq}{}}%
DOI Resolution Documentation. \href{https://www.doi.org/factsheets/DOIProxy.html}{\emph{https://www.doi.org/factsheets/DOIProxy.html}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-xuQKRRx5}{}}%
DOIP and Examples --- Cordra documentation. \href{https://www.cordra.org/documentation/api/doip.html}{\emph{https://www.cordra.org/documentation/api/doip.html}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-p7nwRj8E}{}}%
DOIP API for HTTP Clients --- Cordra documentation. \href{https://www.cordra.org/documentation/api/doip-api-for-http-clients.html}{\emph{https://www.cordra.org/documentation/api/doip-api-for-http-clients.html}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-15EZ2D0Rm}{}}%
draft-bhutton-json-schema-00. \href{https://datatracker.ietf.org/doc/html/draft-bhutton-json-schema-00}{\emph{https://datatracker.ietf.org/doc/html/draft-bhutton-json-schema-00}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-nlH2rm1s}{}}%
draft-ietf-mediaman-suffixes-00 - Media Types with Multiple Suffixes. \href{https://datatracker.ietf.org/doc/draft-ietf-mediaman-suffixes/00/}{\emph{https://datatracker.ietf.org/doc/draft-ietf-mediaman-suffixes/00/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-Amit86Tp}{}}%
draft-ietf-quic-http-34. \href{https://datatracker.ietf.org/doc/html/draft-ietf-quic-http-34}{\emph{https://datatracker.ietf.org/doc/html/draft-ietf-quic-http-34}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-nyrPYHoP}{}}%
draft-kelly-json-hal-08. \href{https://datatracker.ietf.org/doc/html/draft-kelly-json-hal-08}{\emph{https://datatracker.ietf.org/doc/html/draft-kelly-json-hal-08}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-rqjuXkEf}{}}%
Duerst M, Suignard M. 2005. \emph{Internationalized Resource Identifiers (IRIs)}. RFC Editor. DOI: \href{https://doi.org/10.17487/rfc3987}{10.17487/rfc3987}.

\leavevmode\vadjust pre{\hypertarget{ref-xvMAyRAc}{}}%
Dusseault L (ed.). 2007. \emph{HTTP Extensions for Web Distributed Authoring and Versioning (WebDAV)}. RFC Editor. DOI: \href{https://doi.org/10.17487/rfc4918}{10.17487/rfc4918}.

\leavevmode\vadjust pre{\hypertarget{ref-7szz7dwO}{}}%
FAIR Digital Object Framework Documentation. \href{https://fairdigitalobjectframework.org/}{\emph{https://fairdigitalobjectframework.org/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-DMEBhHID}{}}%
FAIR Digital Objects Forum \textbar. \href{https://fairdo.org/}{\emph{https://fairdo.org/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-WzPECPDP}{}}%
Fensel D, Facca FM, Simperl E, Toma I. 2011. \emph{Semantic Web Services}. Springer Berlin Heidelberg. DOI: \href{https://doi.org/10.1007/978-3-642-19193-0}{10.1007/978-3-642-19193-0}.

\leavevmode\vadjust pre{\hypertarget{ref-174AwcFUL}{}}%
Fielding RT. 2000. \href{https://www.ics.uci.edu//~fielding/pubs/dissertation/top.htm}{Architectural styles and the design of network-based software architectures}. \{THESIS\}.\{DOCTORAL\} Thesis. University of California, Irvine.

\leavevmode\vadjust pre{\hypertarget{ref-8HJqcF1Q}{}}%
Fielding R, Nottingham M, Reschke J (eds.). 2022. \emph{HTTP Semantics}. RFC Editor. DOI: \href{https://doi.org/10.17487/rfc9110}{10.17487/rfc9110}.

\leavevmode\vadjust pre{\hypertarget{ref-tIkSXrb5}{}}%
Fielding R, Reschke J (eds.). 2014b. \emph{Hypertext Transfer Protocol (HTTP/1.1): Message Syntax and Routing}. RFC Editor. DOI: \href{https://doi.org/10.17487/rfc7230}{10.17487/rfc7230}.

\leavevmode\vadjust pre{\hypertarget{ref-HE7Ikwwl}{}}%
Fielding R, Reschke J (eds.). 2014a. \emph{Hypertext Transfer Protocol (HTTP/1.1): Semantics and Content}. RFC Editor. DOI: \href{https://doi.org/10.17487/rfc7231}{10.17487/rfc7231}.

\leavevmode\vadjust pre{\hypertarget{ref-11VxBOeZy}{}}%
Fielding RT, Taylor RN, Erenkrantz JR, Gorlick MM, Whitehead J, Khare R, Oreizy P. 2017. Reflections on the REST architectural style and "principled design of the modern web architecture" (impact paper award). In: \emph{Proceedings of the 2017 11th joint meeting on foundations of software engineering - ESEC/FSE 2017}. New York, New York, USA: ACM Press, 4--14. DOI: \href{https://doi.org/10.1145/3106237.3121282}{10.1145/3106237.3121282}.

\leavevmode\vadjust pre{\hypertarget{ref-XmUY0cZB}{}}%
FOAF Vocabulary Specification. \href{http://xmlns.com/foaf/spec/}{\emph{http://xmlns.com/foaf/spec/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-13TcbsZF6}{}}%
Foundation D. 2018. \emph{\href{https://hdl.handle.net/0.DOIP/DOIPV2.0}{Digital object interface protocol specification, version 2.0}}. DONA foundation.

\leavevmode\vadjust pre{\hypertarget{ref-ZeOAPNHa}{}}%
Gayo JEL, Prud'hommeaux E, Boneva I, Kontokostas D. 2017. Validating RDF Data. \emph{Synthesis Lectures on the Semantic Web: Theory and Technology} 7:1--328. DOI: \href{https://doi.org/10.2200/s00786ed1v01y201707wbe016}{10.2200/s00786ed1v01y201707wbe016}.

\leavevmode\vadjust pre{\hypertarget{ref-bMSxa8XW}{}}%
Goble C, Stevens R. 2008. State of the nation in data integration for bioinformatics. \emph{Journal of Biomedical Informatics} 41:687--693. DOI: \href{https://doi.org/10.1016/j.jbi.2008.01.008}{10.1016/j.jbi.2008.01.008}.

\leavevmode\vadjust pre{\hypertarget{ref-IbaebabD}{}}%
Gregorio J, de hOra B (eds.). 2007. \emph{The Atom Publishing Protocol}. RFC Editor. DOI: \href{https://doi.org/10.17487/rfc5023}{10.17487/rfc5023}.

\leavevmode\vadjust pre{\hypertarget{ref-11hxwwuRt}{}}%
Gregorio J, Fielding R, Hadley M, Nottingham M, Orchard D. 2012. \emph{URI Template}. RFC Editor. DOI: \href{https://doi.org/10.17487/rfc6570}{10.17487/rfc6570}.

\leavevmode\vadjust pre{\hypertarget{ref-191ZkYZMt}{}}%
Groth P, Loizou A, Gray AJG, Goble C, Harland L, Pettifer S. 2014. API-centric Linked Data integration: The Open PHACTS Discovery Platform case study. \emph{Journal of Web Semantics} 29:12--18. DOI: \href{https://doi.org/10.1016/j.websem.2014.03.003}{10.1016/j.websem.2014.03.003}.

\leavevmode\vadjust pre{\hypertarget{ref-1GrKTFaK2}{}}%
Group RDAFDMMW. 2020. FAIR data maturity model: Specification and guidelines. \emph{Research Data Alliance}. DOI: \href{https://doi.org/10.15497/rda00050}{10.15497/rda00050}.

\leavevmode\vadjust pre{\hypertarget{ref-1CvYRVhy2}{}}%
Handle.Net Registry. \href{https://www.handle.net/download_hnr.html}{\emph{https://www.handle.net/download\_hnr.html}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-zZTtJruI}{}}%
Hardisty A, Brack P, Goble C, Livermore L, Scott B, Groom Q, Owen S, Soiland-Reyes S. 2022. The Specimen Data Refinery: A Canonical Workflow Framework and FAIR
Digital Object Approach to Speeding up Digital Mobilisation of Natural History
Collections. \emph{Data Intelligence} 4:320--341. DOI: \href{https://doi.org/10.1162/dint_a_00134}{10.1162/dint\_a\_00134}.

\leavevmode\vadjust pre{\hypertarget{ref-GJHYN6xW}{}}%
Hasnain A, Rebholz-Schuhmann D. 2018. Assessing FAIR data principles against the 5-star open data principles. In: Gangemi A, Gentile AL, Nuzzolese AG, Rudolph S, Maleshkova M, Paulheim H, Pan JZ, Alam M eds. \emph{The semantic web: ESWC 2018 satellite events: ESWC 2018 satellite events, heraklion, crete, greece, june 3-7, 2018, revised selected papers}. Lecture notes in computer science. Cham: Springer International Publishing, 469--477. DOI: \href{https://doi.org/10.1007/978-3-319-98192-5_60}{10.1007/978-3-319-98192-5\_60}.

\leavevmode\vadjust pre{\hypertarget{ref-yd2WE9eq}{}}%
Horrocks I, Hendler J (eds.). 2002. \emph{The Semantic Web --- ISWC 2002}. Springer Berlin Heidelberg. DOI: \href{https://doi.org/10.1007/3-540-48005-6}{10.1007/3-540-48005-6}.

\leavevmode\vadjust pre{\hypertarget{ref-14Utscfe4}{}}%
HTML Standard. \href{https://html.spec.whatwg.org/multipage/microdata.html}{\emph{https://html.spec.whatwg.org/multipage/microdata.html}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-FDDmgO6s}{}}%
Hu W, Chen J, Zhang H, Qu Y. 2011. How matchable are four thousand ontologies on the semantic web. In: Antoniou G, Grobelnik M, Simperl E, Parsia B, Plexousakis D, De Leenheer P, Pan J eds. \emph{The semantic web: Research and applications}. Lecture notes in computer science. Berlin, Heidelberg: Springer Berlin Heidelberg, 290--304. DOI: \href{https://doi.org/10.1007/978-3-642-21034-1_20}{10.1007/978-3-642-21034-1\_20}.

\leavevmode\vadjust pre{\hypertarget{ref-13PaKBrZe}{}}%
Hydra W3C Community Group. \href{https://www.hydra-cg.com/}{\emph{https://www.hydra-cg.com/}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-rbG9uRKw}{}}%
Hypertext Style: Cool URIs don't change. \href{https://www.w3.org/Provider/Style/URI}{\emph{https://www.w3.org/Provider/Style/URI}} (accessed 26 May 2022a).

\leavevmode\vadjust pre{\hypertarget{ref-iME5l1kk}{}}%
Hypertext Style: Cool URIs don't change. \href{https://www.w3.org/Provider/Style/URI.html}{\emph{https://www.w3.org/Provider/Style/URI.html}} (accessed 26 May 2022b).

\leavevmode\vadjust pre{\hypertarget{ref-nW4HY8Nq}{}}%
"info" URI Registry (Frozen). \href{https://oclc-research.github.io/infoURI-Frozen/}{\emph{https://oclc-research.github.io/infoURI-Frozen/}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-1GWvyEmUf}{}}%
Iyengar J, Thomson M (eds.). 2021. \emph{QUIC: A UDP-Based Multiplexed and Secure Transport}. RFC Editor. DOI: \href{https://doi.org/10.17487/rfc9000}{10.17487/rfc9000}.

\leavevmode\vadjust pre{\hypertarget{ref-by6CfOmv}{}}%
JSON-LD 1.1. \href{http://www.w3.org/TR/json-ld/}{\emph{http://www.w3.org/TR/json-ld/}} (accessed 26 May 2022b).

\leavevmode\vadjust pre{\hypertarget{ref-X5SWPqSU}{}}%
JSON-LD 1.1. \href{https://www.w3.org/TR/json-ld/}{\emph{https://www.w3.org/TR/json-ld/}} (accessed 26 May 2022a).

\leavevmode\vadjust pre{\hypertarget{ref-HqZ3r6J3}{}}%
Juty N, Le Novere N, Laibe C. 2011. Identifiers.org and MIRIAM Registry: community resources to provide persistent identification. \emph{Nucleic Acids Research} 40:D580--D586. DOI: \href{https://doi.org/10.1093/nar/gkr1097}{10.1093/nar/gkr1097}.

\leavevmode\vadjust pre{\hypertarget{ref-3Uqe3fuK}{}}%
Kahn R, Wilensky R. 1995. \emph{\href{http://www.cnri.reston.va.us/k-w.html}{A framework for distributed digital object services}}. CNRI.

\leavevmode\vadjust pre{\hypertarget{ref-11MnuwJ4l}{}}%
Kahn R, Wilensky R. 2006. A framework for distributed digital object services. \emph{International Journal on Digital Libraries} 6:115--123. DOI: \href{https://doi.org/10.1007/s00799-005-0128-x}{10.1007/s00799-005-0128-x}.

\leavevmode\vadjust pre{\hypertarget{ref-9CdJnsoJ}{}}%
Kamdar MR, Tudorache T, Musen MA. 2017. A systematic analysis of term reuse and term overlap across biomedical ontologies. \emph{Semantic Web} 8:853--871. DOI: \href{https://doi.org/10.3233/sw-160238}{10.3233/sw-160238}.

\leavevmode\vadjust pre{\hypertarget{ref-1B16WfR7q}{}}%
Khare R, Lawrence S. 2000. \emph{Upgrading to TLS Within HTTP/1.1}. RFC Editor. DOI: \href{https://doi.org/10.17487/rfc2817}{10.17487/rfc2817}.

\leavevmode\vadjust pre{\hypertarget{ref-AGHR9oyr}{}}%
Klein M, Van de Sompel H, Sanderson R, Shankar H, Balakireva L, Zhou K, Tobin R. 2014. Scholarly Context Not Found: One in Five Articles Suffers from Reference Rot. \emph{PLoS ONE} 9:e115253. DOI: \href{https://doi.org/10.1371/journal.pone.0115253}{10.1371/journal.pone.0115253}.

\leavevmode\vadjust pre{\hypertarget{ref-o596f2JM}{}}%
Klímek J, Škoda P, Nečaský M. 2019. Survey of tools for linked data consumption. \emph{Satya Widya} 10:665--720. DOI: \href{https://doi.org/10.3233/SW-180316}{10.3233/SW-180316}.

\leavevmode\vadjust pre{\hypertarget{ref-iYh1wFBt}{}}%
Lamprecht A-L, Palmblad M, Ison J, Schwämmle V, Al Manir MS, Altintas I, Baker CJO, Ben Hadj Amor A, Capella-Gutierrez S, Charonyktakis P, Crusoe MR, Gil Y, Goble C, Griffin TJ, Groth P, Ienasescu H, Jagtap P, Kalaš M, Kasalica V, Khanteymoori A, Kuhn T, Mei H, Ménager H, Möller S, Richardson RA, Robert V, Soiland-Reyes S, Stevens R, Szaniszlo S, Verberne S, Verhoeven A, Wolstencroft K. 2021. Perspectives on automated composition of workflows in the life sciences. \emph{F1000Research} 10:897. DOI: \href{https://doi.org/10.12688/f1000research.54159.1}{10.12688/f1000research.54159.1}.

\leavevmode\vadjust pre{\hypertarget{ref-SPd0uEqO}{}}%
Lannom L, Peters-von Gehlen K, Anders I, Pfeil A, Schlemmer A, Trautt Z, Wittenburg P. 2022a. \emph{\href{https://docs.google.com/document/d/1ivvg3C_QWSO9PIQwkKT89xG4fBhNAs7_6b0Dz11EwDg/edit}{FDO configuration types}}. (internal draft): FDO Forum.

\leavevmode\vadjust pre{\hypertarget{ref-ZFzPxCNB}{}}%
Lannom L, Schwardmann U, Blanchi C, Wittenburg P. 2022b. \emph{\href{https://docs.google.com/document/d/1X0hcOVIqP7iYIJf9u-7x3RwcXK8ecsauy0FZg_6-Bg0/edit}{Typing FAIR digital objects}}. (internal draft): FDO Forum.

\leavevmode\vadjust pre{\hypertarget{ref-xVJMmGVr}{}}%
Licenses \& Standards \textbar{} Open Source Initiative. \href{https://opensource.org/licenses}{\emph{https://opensource.org/licenses}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-OGXYtGZ8}{}}%
Linked Data - Design Issues. \href{https://www.w3.org/DesignIssues/LinkedData.html}{\emph{https://www.w3.org/DesignIssues/LinkedData.html}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-zlGYiuWC}{}}%
Linked Data Notifications. \href{https://www.w3.org/TR/ldn/}{\emph{https://www.w3.org/TR/ldn/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-17OHlMRQA}{}}%
Linked Data Platform Working Group. 2015. \emph{\href{https://www.w3.org/TR/2015/REC-ldp-20150226/}{Linked data platform 1.0}}. W3C.

\leavevmode\vadjust pre{\hypertarget{ref-QpDmajc4}{}}%
Linking Across Provenance Bundles. \href{https://www.w3.org/TR/2013/NOTE-prov-links-20130430/}{\emph{https://www.w3.org/TR/2013/NOTE-prov-links-20130430/}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-jrZe6Esu}{}}%
Loo T (ed.). 1970. First International Conference on FAIR Digital Objects. DOI: \href{https://doi.org/10.3897/rio.coll.190}{10.3897/rio.coll.190}.

\leavevmode\vadjust pre{\hypertarget{ref-i9Ang0rM}{}}%
martinekuan.Web API design best practices - Azure Architecture Center. \href{https://learn.microsoft.com/en-us/azure/architecture/best-practices/api-design}{\emph{https://learn.microsoft.com/en-us/azure/architecture/best-practices/api-design}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-PL512B0X}{}}%
May W. 2017. \emph{HTTP Live Streaming}. RFC Editor. DOI: \href{https://doi.org/10.17487/rfc8216}{10.17487/rfc8216}.

\leavevmode\vadjust pre{\hypertarget{ref-5V3GqSzR}{}}%
Meroño-Peñuela A, Lisena P, Martínez-Ortiz C. 2021b. Web data apis over SPARQL. In: \emph{Web data apis for knowledge graphs: Easing access to semantic data for application developers}. Synthesis lectures on data, semantics, and knowledge. Cham: Springer International Publishing, 27--38. DOI: \href{https://doi.org/10.1007/978-3-031-01917-3_3}{10.1007/978-3-031-01917-3\_3}.

\leavevmode\vadjust pre{\hypertarget{ref-99MS5xE0}{}}%
Meroño-Peñuela A, Lisena P, Martínez-Ortiz C. 2021a. Conclusion and future challenges. In: \emph{Web data apis for knowledge graphs: Easing access to semantic data for application developers}. Synthesis lectures on data, semantics, and knowledge. Cham: Springer International Publishing, 77--80. DOI: \href{https://doi.org/10.1007/978-3-031-01917-3_7}{10.1007/978-3-031-01917-3\_7}.

\leavevmode\vadjust pre{\hypertarget{ref-tZOPTKnJ}{}}%
Mika P. 2015. On Schema.org and Why It Matters for the Web. \emph{IEEE Internet Computing} 19:52--55. DOI: \href{https://doi.org/10.1109/mic.2015.81}{10.1109/mic.2015.81}.

\leavevmode\vadjust pre{\hypertarget{ref-u8Cy0psL}{}}%
Mons B, Neylon C, Velterop J, Dumontier M, Silva Santos LOB da, Wilkinson MD. 2017. Cloudy, increasingly FAIR; revisiting the FAIR data guiding principles for the european open science cloud. \emph{Information Services \& Use} 37:49--56. DOI: \href{https://doi.org/10.3233/\%7BISU\%7D-170824}{10.3233/\{ISU\}-170824}.

\leavevmode\vadjust pre{\hypertarget{ref-1Gsq23e51}{}}%
NCBO BioPortal. \href{https://bioportal.bioontology.org/ontologies}{\emph{https://bioportal.bioontology.org/ontologies}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-J2H8yssV}{}}%
Neumann A, Laranjeiro N, Bernardino J. 2021. An analysis of public REST web service apis. \emph{IEEE Transactions on Services Computing} 14:957--970. DOI: \href{https://doi.org/10.1109/TSC.2018.2847344}{10.1109/TSC.2018.2847344}.

\leavevmode\vadjust pre{\hypertarget{ref-ozdWB3O7}{}}%
Nottingham M. 2017. \emph{Web Linking}. RFC Editor. DOI: \href{https://doi.org/10.17487/rfc8288}{10.17487/rfc8288}.

\leavevmode\vadjust pre{\hypertarget{ref-k0AfCGzw}{}}%
OpenAPI Specification v3.1.0 \textbar{} Introduction, Definitions, \& More. \href{https://spec.openapis.org/oas/v3.1.0.html}{\emph{https://spec.openapis.org/oas/v3.1.0.html}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-ID1pTGry}{}}%
ORE Specification - Abstract Data Model. \href{http://www.openarchives.org/ore/1.0/datamodel\#Proxies}{\emph{http://www.openarchives.org/ore/1.0/datamodel\#Proxies}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-gtGuSevK}{}}%
ORE User Guide - Primer. \href{http://www.openarchives.org/ore/1.0/primer}{\emph{http://www.openarchives.org/ore/1.0/primer}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-1p4IWJpI}{}}%
OWL 2 Web Ontology Language Document Overview (Second Edition). \href{http://www.w3.org/TR/owl2-overview/}{\emph{http://www.w3.org/TR/owl2-overview/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-eS6a6tlz}{}}%
Page KR, De Roure DC, Martinez K. 2011. REST and Linked Data. In: \emph{Proceedings of the Second International Workshop on RESTful Design - WS-REST '11}. ACM Press,. DOI: \href{https://doi.org/10.1145/1967428.1967435}{10.1145/1967428.1967435}.

\leavevmode\vadjust pre{\hypertarget{ref-5SuwLNdE}{}}%
Polleres A, Kamdar MR, Fernández JD, Tudorache T, Musen MA. 2020. A more decentralized vision for linked data. \emph{Satya Widya} 11:101--113. DOI: \href{https://doi.org/10.3233/SW-190380}{10.3233/SW-190380}.

\leavevmode\vadjust pre{\hypertarget{ref-9T4j3N4e}{}}%
PROV-O: The PROV Ontology. \href{https://www.w3.org/TR/2013/REC-prov-o-20130430/}{\emph{https://www.w3.org/TR/2013/REC-prov-o-20130430/}} (accessed 24 January 2023a).

\leavevmode\vadjust pre{\hypertarget{ref-ZUgwvAHE}{}}%
PROV-O: The PROV Ontology. \href{https://www.w3.org/TR/prov-o/\#specializationOf}{\emph{https://www.w3.org/TR/prov-o/\#specializationOf}} (accessed 24 January 2023b).

\leavevmode\vadjust pre{\hypertarget{ref-16WGmCih8}{}}%
RDF 1.1 Primer. \href{http://www.w3.org/TR/rdf11-primer/}{\emph{http://www.w3.org/TR/rdf11-primer/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-ZwAcGQKY}{}}%
RDF Schema 1.1. \href{http://www.w3.org/TR/rdf-schema/}{\emph{http://www.w3.org/TR/rdf-schema/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-FKIF6ApB}{}}%
RDFa 1.1 Primer - Third Edition. \href{http://www.w3.org/TR/rdfa-primer/}{\emph{http://www.w3.org/TR/rdfa-primer/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-16wvFH4nh}{}}%
Resource Description Framework (RDF) Model and Syntax Specification. \href{https://www.w3.org/TR/1999/REC-rdf-syntax-19990222/}{\emph{https://www.w3.org/TR/1999/REC-rdf-syntax-19990222/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-eS5y9TRh}{}}%
ResourceSync Framework Specification - Table of Contents. \href{http://www.openarchives.org/rs/toc}{\emph{http://www.openarchives.org/rs/toc}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-F8VZ86hu}{}}%
Sandvine.Global Internet Phenomena Report 2022. \href{https://www.sandvine.com/global-internet-phenomena-report-2022}{\emph{https://www.sandvine.com/global-internet-phenomena-report-2022}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-rnzJ4khD}{}}%
Sauermann L, Cyganiak R, Völkel M. 2011. \emph{Cool URIs for the semantic web}. Universität des Saarlandes. DOI: \href{https://doi.org/10.22028/d291-25086}{10.22028/d291-25086}.

\leavevmode\vadjust pre{\hypertarget{ref-10pzKCEUE}{}}%
Schema.org - Schema.org. \href{https://schema.org/}{\emph{https://schema.org/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-dKAekUjL}{}}%
Schema.org Actions - schema.org. \href{https://schema.org/docs/actions.html}{\emph{https://schema.org/docs/actions.html}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-IHLT6hye}{}}%
Schultes E, Wittenburg P. 2019. FAIR principles and digital objects: Accelerating convergence on a data infrastructure. In: Manolopoulos Y, Stupnikov S eds. \emph{Data analytics and management in data intensive domains: 20th international conference, DAMDID/RCDL 2018, moscow, russia, october 9--12, 2018, revised selected papers}. Communications in computer and information science. Cham: Springer International Publishing, 3--16. DOI: \href{https://doi.org/\%7B10.1007/978-3-030-23584-0_1\%7D}{\{10.1007/978-3-030-23584-0\_1\}}.

\leavevmode\vadjust pre{\hypertarget{ref-1CqIZ47pu}{}}%
Schwardmann U, Kálmán T. 2022. Two Examples on How FDO Types can Support Machine and Human Readability. \emph{Research Ideas and Outcomes} 8. DOI: \href{https://doi.org/10.3897/rio.8.e96014}{10.3897/rio.8.e96014}.

\leavevmode\vadjust pre{\hypertarget{ref-kYjzRJ0H}{}}%
Semantic Web - XML2000 - slide "Architecture". \href{https://www.w3.org/2000/Talks/1206-xml2k-tbl/slide10-0.html}{\emph{https://www.w3.org/2000/Talks/1206-xml2k-tbl/slide10-0.html}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-YUzogaV}{}}%
Shape Expressions (ShEx) 2.1 Primer. \href{http://shex.io/shex-primer/}{\emph{http://shex.io/shex-primer/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-GDrm9dgX}{}}%
Shapes Constraint Language (SHACL). \href{https://www.w3.org/TR/shacl/}{\emph{https://www.w3.org/TR/shacl/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-15gQDya5B}{}}%
SKOS Simple Knowledge Organization System Primer. \href{http://www.w3.org/TR/skos-primer/}{\emph{http://www.w3.org/TR/skos-primer/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-CNP39tah}{}}%
Smith B, Ashburner M, Rosse C, Bard J, Bug W, Ceusters W, Goldberg LJ, Eilbeck K, Ireland A, Mungall CJ, Leontis N, Rocca-Serra P, Ruttenberg A, Sansone S-A, Scheuermann RH, Shah N, Whetzel PL, Lewis S. 2007. The OBO Foundry: coordinated evolution of ontologies to support biomedical data integration. \emph{Nature Biotechnology} 25:1251--1255. DOI: \href{https://doi.org/10.1038/nbt1346}{10.1038/nbt1346}.

\leavevmode\vadjust pre{\hypertarget{ref-xww8fb1B}{}}%
SPARQL 1.1 Overview. \href{http://www.w3.org/TR/sparql11-overview/}{\emph{http://www.w3.org/TR/sparql11-overview/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-8ZRGFa6y}{}}%
Stallings W. 1990. \emph{Handbook of computer-communications standards: The open systems (OSI) model and OSI-related standards}. Carmel, IN, USA: Sams.

\leavevmode\vadjust pre{\hypertarget{ref-NxvhwSmG}{}}%
Stanczyk SK. 1987. Process modelling for information system description. \emph{The Open University}. DOI: \href{https://doi.org/10.21954/ou.ro.0000f821}{10.21954/ou.ro.0000f821}.

\leavevmode\vadjust pre{\hypertarget{ref-uxWjmlio}{}}%
Stefi A. 2015. Do Developers Make Unbiased Decisions? - The Effect of Mindfulness and Not-Invented-Here Bias on the Adoption of Software Components. DOI: \href{https://doi.org/10.18151/7217489}{10.18151/7217489}.

\leavevmode\vadjust pre{\hypertarget{ref-v0WwFHgT}{}}%
Stefi A, Hess T. 2015. To develop or to reuse? Two perspectives on external reuse in software projects. In: Fernandes JM, Machado RJ, Wnuk K eds. \emph{Software business}. Lecture notes in business information processing. Cham: Springer International Publishing, 192--206. DOI: \href{https://doi.org/10.1007/978-3-319-19593-3_18}{10.1007/978-3-319-19593-3\_18}.

\leavevmode\vadjust pre{\hypertarget{ref-yygVPoL0}{}}%
Strawn G, Wittenburg P, Soiland-Reyes St, Peters K, Lannom L, Schwardmann U, Blanch Ch. 2022. \emph{\href{https://docs.google.com/document/d/1aGA-TBr4XpORhMPtnf_--Nb4FYJccgeSvGmGh68jNws/edit}{FDO Forum FDO requirement specifications}}. (internal draft): FDO Forum.

\leavevmode\vadjust pre{\hypertarget{ref-2gXY2Nwh}{}}%
Sun S, Lannom L, Boesch B. 2003. \emph{Handle System Overview}. RFC Editor. DOI: \href{https://doi.org/10.17487/rfc3650}{10.17487/rfc3650}.

\leavevmode\vadjust pre{\hypertarget{ref-I9Kx0Hjy}{}}%
Sun S, Reilly S, Lannom L, Petrone J. 2003. \emph{Handle System Protocol (ver 2.1) Specification}. RFC Editor. DOI: \href{https://doi.org/10.17487/rfc3652}{10.17487/rfc3652}.

\leavevmode\vadjust pre{\hypertarget{ref-7DGllKzG}{}}%
SWORD 3.0 Specification. \href{https://swordapp.github.io/swordv3/swordv3.html}{\emph{https://swordapp.github.io/swordv3/swordv3.html}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-voDQ6diM}{}}%
The ARK Identifier Scheme. \href{https://datatracker.ietf.org/doc/id/draft-kunze-ark.html}{\emph{https://datatracker.ietf.org/doc/id/draft-kunze-ark.html}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-hxtgCGjY}{}}%
The Modern Standards Paradigm - Five Key Principles. \href{https://open-stand.org/about-us/principles/}{\emph{https://open-stand.org/about-us/principles/}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-oN9D5LkD}{}}%
The Open Graph protocol. \href{https://ogp.me/}{\emph{https://ogp.me/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-S5BI6Mwx}{}}%
Thornton K, Solbrig H, Stupp GS, Labra Gayo JE, Mietchen D, Prud E, Waagmeester A. 2019. Using shape expressions (ShEx) to share RDF data models and to guide curation with rigorous validation. In: Hitzler P, Fernández M, Janowicz K, Zaveri A, Gray AJG, Lopez V, Haller A, Hammar K eds. \emph{The semantic web: 16th international conference, ESWC 2019, portorož, slovenia, june 2--6, 2019, proceedings}. Lecture notes in computer science. Cham: Springer International Publishing, 606--620. DOI: \href{https://doi.org/10.1007/978-3-030-21348-0_39}{10.1007/978-3-030-21348-0\_39}.

\leavevmode\vadjust pre{\hypertarget{ref-1DalZuKct}{}}%
Tirmizi S, Aitken S, Moreira DA, Mungall C, Sequeda J, Shah NH, Miranker DP. 2011. Mapping between the OBO and OWL ontology languages. \emph{Journal of Biomedical Semantics} 2:S3. DOI: \href{https://doi.org/10.1186/2041-1480-2-s1-s3}{10.1186/2041-1480-2-s1-s3}.

\leavevmode\vadjust pre{\hypertarget{ref-DNTCUjpC}{}}%
Turcoane O. 2014. Linked data, JSON-LD and the semantics of cultural and scientific heritage. \emph{Digital Presentation and Preservation of Cultural and Scientific Heritage} 4:95--105. DOI: \href{https://doi.org/10.55630/dipp.2014.4.11}{10.55630/dipp.2014.4.11}.

\leavevmode\vadjust pre{\hypertarget{ref-10HA0mMBx}{}}%
Unknown.Introducing 'Role'. \href{http://blog.schema.org/2014/06/introducing-role.html}{\emph{http://blog.schema.org/2014/06/introducing-role.html}} (accessed 24 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-j3SqQLBR}{}}%
Usage Statistics of JSON-LD for Websites, May 2022. \href{https://w3techs.com/technologies/details/da-jsonld}{\emph{https://w3techs.com/technologies/details/da-jsonld}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-snykkm7R}{}}%
Van de Sompel H, Klein M, Jones S, Nelson ML, Warner S, Devaraju A, Huber R, Steinhoff W, Tykhonov V, Boruta L, Meijers E, Soiland-Reyes S, Wilkinson M. 2022.FAIR Signposting Profile. \href{https://signposting.org/FAIR/}{\emph{https://signposting.org/FAIR/}} (accessed 5 January 2023).

\leavevmode\vadjust pre{\hypertarget{ref-DNjcMi4a}{}}%
Van de Sompel H, Nelson M, Sanderson R. 2013. \emph{HTTP Framework for Time-Based Access to Resource States -\/- Memento}. RFC Editor. DOI: \href{https://doi.org/10.17487/rfc7089}{10.17487/rfc7089}.

\leavevmode\vadjust pre{\hypertarget{ref-jHwttlTs}{}}%
Verborgh R, Vander Sande M. 2020. The semantic web identity crisis: In search of the trivialities that never were. \emph{Satya Widya} 11:19--27. DOI: \href{https://doi.org/10.3233/SW-190372}{10.3233/SW-190372}.

\leavevmode\vadjust pre{\hypertarget{ref-11oVTMsuZ}{}}%
W3C XML Schema Definition Language (XSD) 1.1 Part 1: Structures. \href{http://www.w3.org/TR/xmlschema11-1/}{\emph{http://www.w3.org/TR/xmlschema11-1/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-LmudecyN}{}}%
Web Services Description Language (WSDL) Version 2.0 Part 0: Primer. \href{http://www.w3.org/TR/wsdl20-primer/}{\emph{http://www.w3.org/TR/wsdl20-primer/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-qgqADk45}{}}%
WebSockets Standard. \href{https://websockets.spec.whatwg.org/}{\emph{https://websockets.spec.whatwg.org/}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-lCoshHXg}{}}%
Weigel T, Plale B, Parsons M, Zhou G, Luo Y, Schwardmann U, Quick R, Hellström M, Kurakawa K. 2018. RDA Recommendation on PID Kernel Information. \emph{Research Data Alliance}. DOI: \href{https://doi.org/10.15497/rda00031}{10.15497/rda00031}.

\leavevmode\vadjust pre{\hypertarget{ref-iJeys0T5}{}}%
Weiland C, Islam S, Broder D, Anders I, Wittenburg P. 2022a. \emph{\href{https://docs.google.com/document/d/1GHFPAUGpNvYaxctkx-CpvY1vKf_aGZpSlWGOWvyXSiQ/edit}{FDO machine actionability}}. (internal draft): FDO Forum.

\leavevmode\vadjust pre{\hypertarget{ref-TQdku4YF}{}}%
Weiland C, Schwardmann U, Wittenburg P, Kirkpatrick C, Hanisch R, Trautt Z. 2022b. \emph{\href{https://docs.google.com/document/d/1lPNBBROjEoZ6fTfrtdqcMa3Q2G27PoC_/edit}{FDO Forum Document Standards}}. (internal draft): FDO Forum.

\leavevmode\vadjust pre{\hypertarget{ref-9jaVzCbL}{}}%
Wieczorek J, Bloom D, Guralnick R, Blum S, Döring M, Giovanni R, Robertson T, Vieglais D. 2012. Darwin Core: An Evolving Community-Developed Biodiversity Data Standard. \emph{PLoS ONE} 7:e29715. DOI: \href{https://doi.org/10.1371/journal.pone.0029715}{10.1371/journal.pone.0029715}.

\leavevmode\vadjust pre{\hypertarget{ref-nCzRO0pK}{}}%
Wilde E. 2013. \emph{The 'profile' Link Relation Type}. RFC Editor. DOI: \href{https://doi.org/10.17487/rfc6906}{10.17487/rfc6906}.

\leavevmode\vadjust pre{\hypertarget{ref-6DjakjNS}{}}%
Wilkinson MD, Dumontier M, Aalbersberg IjJ, Appleton G, Axton M, Baak A, Blomberg N, Boiten J-W, da Silva Santos LB, Bourne PE, Bouwman J, Brookes AJ, Clark T, Crosas M, Dillo I, Dumon O, Edmunds S, Evelo CT, Finkers R, Gonzalez-Beltran A, Gray AJG, Groth P, Goble C, Grethe JS, Heringa J, 't Hoen PAC, Hooft R, Kuhn T, Kok R, Kok J, Lusher SJ, Martone ME, Mons A, Packer AL, Persson B, Rocca-Serra P, Roos M, van Schaik R, Sansone S-A, Schultes E, Sengstag T, Slater T, Strawn G, Swertz MA, Thompson M, van der Lei J, van Mulligen E, Velterop J, Waagmeester A, Wittenburg P, Wolstencroft K, Zhao J, Mons B. 2016. The FAIR Guiding Principles for scientific data management and stewardship. \emph{Scientific Data} 3. DOI: \href{https://doi.org/10.1038/sdata.2016.18}{10.1038/sdata.2016.18}.

\leavevmode\vadjust pre{\hypertarget{ref-Pa67pUtR}{}}%
Wilkinson SR, Eisenhauer G, Kapadia AJ, Knight K, Logan J, Widener P, Wolf M. 2022. F*** workflows: when parts of FAIR are missing. \emph{arXiv}. DOI: \href{https://doi.org/10.48550/arxiv.2209.09022}{10.48550/arxiv.2209.09022}.

\leavevmode\vadjust pre{\hypertarget{ref-Qlrlr2TZ}{}}%
Williams AJ, Harland L, Groth P, Pettifer S, Chichester C, Willighagen EL, Evelo CT, Blomberg N, Ecker G, Goble C, Mons B. 2012. Open PHACTS: semantic interoperability for drug discovery. \emph{Drug Discovery Today} 17:1188--1198. DOI: \href{https://doi.org/10.1016/j.drudis.2012.05.016}{10.1016/j.drudis.2012.05.016}.

\leavevmode\vadjust pre{\hypertarget{ref-jRWurmQc}{}}%
Wittenburg P, Anders I, Blanchi C, Buurman M, Goble C, Grieb J, Hardisty A, Islam S, Jejkal T, Kálmán T, Kirkpatrick C, Lannom L, Lauer T, Manepalli G, Peters-von Gehlen K, Pfeil A, Quick R, Sanden M van de, Schwardmann U, Soiland-Reyes S, Stotzka R, Trautt Z, Van Uytvanck D, Weiland C, Wieder P. 2022. FAIR digital object demonstrators 2021. \emph{Zenodo}. DOI: \href{https://doi.org/10.5281/zenodo.5872645}{10.5281/zenodo.5872645}.

\leavevmode\vadjust pre{\hypertarget{ref-tz0P3DTC}{}}%
Wittenburg P, Strawn G, Mons B, Bonino L, Schultes E. 2019. Digital objects as drivers towards convergence in data infrastructures. \emph{https://b2share.eudat.eu}. DOI: \href{https://doi.org/10.23728/b2share.b605d85809ca45679b110719b6c6cb11}{10.23728/b2share.b605d85809ca45679b110719b6c6cb11}.

\leavevmode\vadjust pre{\hypertarget{ref-39K3yHT3}{}}%
Wolstencroft K, Haines R, Fellows D, Williams A, Withers D, Owen S, Soiland-Reyes S, Dunlop I, Nenadic A, Fisher P, Bhagat J, Belhajjame K, Bacall F, Hardisty A, Nieva de la Hidalga A, Balcazar Vargas MP, Sufi S, Goble C. 2013. The Taverna workflow suite: designing and executing workflows of Web Services on the desktop, web or in the cloud. \emph{Nucleic Acids Research} 41:W557--W561. DOI: \href{https://doi.org/10.1093/nar/gkt328}{10.1093/nar/gkt328}.

\leavevmode\vadjust pre{\hypertarget{ref-MhsCt4P3}{}}%
Wolstencroft K, Owen S, Horridge M, Krebs O, Mueller W, Snoep JL, du Preez F, Goble C. 2011. RightField: embedding ontology annotation in spreadsheets. \emph{Bioinformatics} 27:2021--2022. DOI: \href{https://doi.org/10.1093/bioinformatics/btr312}{10.1093/bioinformatics/btr312}.

\leavevmode\vadjust pre{\hypertarget{ref-15yGpJ0wh}{}}%
Working Group FDO-TSIG. 2022a. \emph{\href{https://docs.google.com/document/d/1me0L8C5yDe39cYP1Sxud4Y10hxhphOimLB-K-KgHQUk/edit}{DOIP endorsement request}}. (internal draft): FDO Forum.

\leavevmode\vadjust pre{\hypertarget{ref-16W9AycSW}{}}%
Working Group FDO-TSIG. 2022b. \emph{\href{https://docs.google.com/document/d/1VJTYHxLXIbjHSoFnMEkKaTbpTz9CagoIJyoSafkVh_I/edit}{FDO -- granularity, versioning, mutability}}. (internal draft): FDO Forum.

\leavevmode\vadjust pre{\hypertarget{ref-Fi62cJAo}{}}%
Working Group FDO-TSIG, Working Group FDO-BIG. 2022. \emph{\href{https://docs.google.com/document/d/11CuSoNOpg3vYaoHqi4Yg6ncP3qOclJIC/edit}{FDO -- kernel attributes \& metadata}}. (internal draft): FDO Forum.

\leavevmode\vadjust pre{\hypertarget{ref-103Hw8H43}{}}%
X.1255 : Framework for discovery of identity management information. \href{https://www.itu.int/rec/T-REC-X.1255-201309-I}{\emph{https://www.itu.int/rec/T-REC-X.1255-201309-I}} (accessed 26 May 2022).

\leavevmode\vadjust pre{\hypertarget{ref-hRzcHhPD}{}}%
Zarras A. 2004. A Comparison Framework for Middleware Infrastructures. \emph{The Journal of Object Technology} 3:103. DOI: \href{https://doi.org/10.5381/jot.2004.3.5.a2}{10.5381/jot.2004.3.5.a2}.

\end{CSLReferences}