### 2023-03-21 latex/table5.tex

\begin{landscape}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2}}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1}}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2}}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2}}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2}}@{}}
\caption{Assessing RDA's FAIR Data Maturity Model \autocite{1GrKTFaK2,UzQhqk0M} (first 2 columns) against the FDO guidelines \autocite{RwvirqWg}, FDO implemented with the protocol DOIPv2 \autocite{13TcbsZF6}, Linked Data Platform (LDP) \autocite{7szz7dwO} and examples from Linked Data practices in general. (--- indicates \emph{Unspecified}, may be possible with additional conventions)
\label{tbl:fair-data-maturity-model}}\tabularnewline
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
FAIR ID
\end{minipage}
  & \begin{minipage}[b]{\linewidth}\raggedright
Indicator
\end{minipage}
  & \begin{minipage}[b]{\linewidth}\centering
FDO guidelines
\end{minipage}
  & \begin{minipage}[b]{\linewidth}\centering
FDO/DOIP
\end{minipage}
  & \begin{minipage}[b]{\linewidth}\centering
FDO/LDP
\end{minipage}
  & \begin{minipage}[b]{\linewidth}\centering
Linked Data examples
\end{minipage} \\
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{\linewidth}\raggedright
FAIR ID
\end{minipage}
  & \begin{minipage}[b]{\linewidth}\raggedright
Indicator
\end{minipage}
  & \begin{minipage}[b]{\linewidth}\centering
FDO guidelines
\end{minipage}
  & \begin{minipage}[b]{\linewidth}\centering
FDO/DOIP
\end{minipage}
  & \begin{minipage}[b]{\linewidth}\centering
FDO/LDP
\end{minipage}
  & \begin{minipage}[b]{\linewidth}\centering
Linked Data examples
\end{minipage} \\
\midrule
\endhead
RDA-F1-01M
  & Metadata is identified by a persistent identifier
  & FDOF4
  & Optional \emph{Metadata FDO} w/separate PID
  & Content-negotiation to URL, not required to be PID
  & Metadata typically don't have own PID \\
RDA-F1-01D
  & Data is identified by a persistent identifier
  & FDOF1
  & PIDs required (FDOF1). Handle, DOI.
  & FDOF-IR (Identifier Record). PID can be any URI
  & ``Cool'' URIs \autocite{berners-lee-cool-uris}, PURL services incl.~\texttt{purl.org}, \texttt{w3id.org} \\
RDA-F1-02M
  & Metadata is identified by a globally unique identifier
  & FDOR4 FDOF8
  & Optional \emph{Metadata FDO}, unspecified how to indicate
  & Content-negotiation to URL
  & Not required, content-negotiation can redirect to URL or \texttt{Content-Location}. FAIR Signposting. \\
RDA-F1-02D
  & Data is identified by a globally unique identifier
  & FDOF1
  & All FDOs have PIDs (FDOR1), DOIP uses Handle system
  & FDOF-IR (Identifier Record)
  & Always accessed by URL \\
RDA-F2-01M
  & Rich metadata is provided to allow discovery
  & FDOF2 FDOF4 FDOF8 FDOF9
  & FDO has key-value metadata. Unclear how to link to additional metadata.
  & FDOF-IR links to multiple metadata records
  & RDF-based metadata by content negotiation or FAIR Signposting. Embedded in landing page (RDFa). \\
RDA-F3-01M
  & Metadata includes the identifier for the data
  & ---
  & \texttt{id} and \texttt{type} are required metadata elements PIDs, also implicit as requests must use PID
  & PID only required in FDOF-IR record.
  & PID inclusion typical, but often inconsistent (e.g.~\texttt{www.example.com} vs \texttt{example.com}) or missing (use of \texttt{\textless{}\textgreater{}} as \emph{this} subject) \\
RDA-F4-01M
  & Metadata is offered in such a way that it can be harvested and indexed
  & FDOF10
  & No, registries not required (except Data Type Registries). Handle registry only searchable by PID.
  & Not specified
  & Not specified, several registries/catalogues for vocabularies/types (e.g. \autocite{NCBOBioPortal}). Indexing by search engines if exposing HTML w/schema.org. \\
RDA-A1-01M
  & Metadata contains information to enable the user to get access to the data
  & FDOF3 FDOF6
  & Directly by DOIP, but not included in FDO metadata. \texttt{handle.net} HTTP resolution may redirect to landing page
  & Any property can point to URIs, but unclear if it is data
  & Common with clickable ``follow your nose'' URLs \\
RDA-A1-02M
  & Metadata can be accessed manually (i.e.~with human intervention)
  & ---
  & (Cordra HTML landing page from \texttt{handle.net} URIs)
  & Optional content-negotiation, e.g.~by Apache Marmotta, OpenLink Virtuoso
  & HTTP content-negotiation to HTML is common \\
RDA-A1-02D
  & Data can be accessed manually (i.e.~with human intervention)
  & ---
  & (Cordra HTML landing page from \texttt{handle.net} URIs)
  & Optional content-negotiation
  & Direct download, HTML landing pages common for DOIs \\
RDA-A1-03M
  & Metadata identifier resolves to a metadata record
  & FDOF8+FDOF2
  & ---
  & ---
  & \texttt{Content-Location} or HTTP redirection may indicate metadata URI \\
RDA-A1-03D
  & Data identifier resolves to a digital object
  & FDOF2
  & Required, but frequently not directly resolvable
  & Recommended, but any URI acceptable
  & Resolvable HTTP/HTTPS URIs are most common, now infrequent URNs are not directly resolvable \\
RDA-A1-04M
  & Metadata is accessed through standardised protocol
  & G9 FDOF3
  & Retrievable from PID (FDOF3). Informal DOIP standard maintained by DONA Foundation
  & LDP standard maintained by W3C, HTTP standards maintained by IETF, FDO components resolved by informal proposals (custom vocabulary, extra HTTP methods) or HTTP content negotiation)
  & Formal HTTP standards maintained by IETF, HTTP content negotiation, informal FAIR Signposting \\
RDA-A1-04D
  & Data is accessible through standardised protocol
  & G9
  & (see above)
  & HTTP \autocite{rfc9110}
  & HTTP/HTTPS, FTP (now less common), GridFTP \autocite{allcockGlobusStripedGridFTP} (for large data), ARK \autocite{ARKIdentifierScheme} \\
RDA-A1-05D
  & Data can be accessed automatically (i.e.~by a computer program)
  & G4 FDOF3 FDOF6
  & Required, but few client libraries
  & & Ubiquitous, hundreds of HTTP libraries \\
RDA-A1.1-01M
  & Metadata is accessible through a free access protocol    
  & G1 G8 G9
  & Partially realised: Handle system is open\footnote{
        The \texttt{Handle.net} system was previously covered by software patent \href{https://patents.google.com/patent/US6135646A/en}{US6135646A} which \href{https://circleid.com/posts/20161025_selling_dona_snake_oil_at_the_itu\#11461}{expired} in 2013.} 
    protocol \autocite{rfc3652}. One server implementation \autocite{HandleNetRegistry}, free\footnote{
        The \href{http://www.handle.net/HNRj/HNR-9-License.pdf}{Handle.net public license} is not OSI-approved \autocite{LicensesStandardsOpen}  as an open source license -- it includes usage restrictions and requires Service Agreements. It is not a DOIP requirement to host a local Handle instance, e.g.~EOSC provides the \href{https://sp.eudat.eu/catalog/resources/fc6b2d30-09cd-4c25-b71a-7bc6de77910c}{B2HANDLE} service for acquiring Handle prefixes.}. 
    One DOIPv2 implementation (\href{https://www.cordra.org/}{CORDRA}): free under BSD-like license (not recognised as Open Source).    
  & LDP is open W3C recommendation \autocite{w3-ldp}. \href{https://www.w3.org/wiki/LDP_Implementations}{Multiple LDP implementations}.    
  & DNS, HTTP, TLS, RDF standards are open, free and universal, large number of Open Source clients and \href{https://en.wikipedia.org/wiki/Comparison_of_web_server_software}{servers}. \\
RDA-A1.1-01D
  & Data is accessible through a free access protocol
  & G9
  & (see above)
  & URI, DNS, HTTP, TLS
  & URI, DNS, HTTP, TLS. Non-free DRM may be used (e.g.~subscription video streaming) \\
RDA-A1.2-01D
  & Data is accessible through an access protocol that supports authentication and authorisation
  & (FDOR9)
  & TLS certificates, \texttt{authentication} field (details unspecified)
  & Implied
  & HTTP authentication, TLS certificates \\
RDA-A2-01M\label{RDA-A2-01M}
  & Metadata is guaranteed to remain available after data is no longer available
  & FDOF12
  & ---
  & Unspecified, however FDOF-IR links to separate metadata records
  & --- \\
RDA-I1-01M
  & Metadata uses knowledge representation expressed in standardised format
  & FDOF8
  & Required, but not currently defined
  & ---
  & Always implied by use of RDF syntaxes. \\
RDA-I1-01D
  & Data uses knowledge representation expressed in standardised format
  & ---
  & ---
  & ---
  & Common (e.g.~HDF5, JSON, XML), yet common scientific data formats frequently not standardised \\
RDA-I1-02M
  & Metadata uses machine-understandable knowledge representation
  & FDOF8
  & Required
  & Optional RDF metadata with any vocabulary
  & Always implied by use of RDF syntaxes. \\
RDA-I1-02D
  & Data uses machine-understandable knowledge representation
  & G4 G7 FDOR2
  & No requirements on binary data formats
  & Only indirectly, \href{https://www.w3.org/TR/ldp/\#dfn-linked-data-platform-basic-container}{LDP Basic Container} reference only information resources
  & Common, specially for scientific data formats \\
RDA-I2-01M
  & Metadata uses FAIR-compliant vocabularies
  & G3 FDOF10
  & Informally required
  & Unspecified, implied by use of RDF?
  & FAIR practices for LD vocabularies increasingly common, sometimes inconsistent (e.g.~PURLs that don't resolve) or incomplete (e.g.~unknown license) \\
RDA-I2-01D
  & Data uses FAIR-compliant vocabularies
  & ---
  & ---
  & ---
  & Uncommon, except for some XML and RDF-embedding formats, e.g.~Extensible Metadata Platform (XMP) \autocite{iso16684} \\
RDA-I3-01M
  & Metadata includes references to other metadata
  & FDOR8
  & Implied (attributes to PIDs), currently unspecified if given attribute is value or reference
  & ---
  & By definition (Linked Data reference existing URIs \autocite{DataW3C}), \texttt{rdfs:seeAlso}, FAIR signposting \autocite{vandesompelFAIRSignpostingProfile2022} \texttt{describedby} \\
RDA-I3-01D
  & Data includes references to other data
  & G6 FDOR3 FDOR11
  & ---
  & ---
  & URL hyperlinks common in several formats (HTML, PDF, JSON, XML). \\
RDA-I3-02M
  & Metadata includes references to other data
  & G6 FDOR3 FDOR8
  & Implied from custom FDO type's attribute
  & LDP Direct Container members can be any resources
  & URI objects are frequently data references, may be indirect via PID \\
RDA-I3-02D
  & Data includes qualified references to other data
  & FDOR3 FDOR11
  & Only indirectly through FDO metadata
  & Indirectly through LDP membership
  & Uncommon: Link relations, FAIR Signposting \\
RDA-I3-03M
  & Metadata includes qualified references to other metadata
  & (FDOR3)
  & Qualification by attribute keys defined per FDO Type
  & \href{https://www.w3.org/TR/ldp/\#dfn-linked-data-platform-direct-container}{LDP Direct Container}
  & Qualifications by property, PROV bundles \autocite{w3-prov-links}, \href{https://schema.org/Role}{schema.org/Role} \\
RDA-I3-04M
  & Metadata include qualified references to other data
  & (FDOR3)
  & Qualification by attribute keys defined per FDO type
  & \href{https://www.w3.org/TR/ldp/\#dfn-linked-data-platform-indirect-container}{LDP Indirect Container}
  & Qualifications by property, n-ary indirection (schema.org Role \autocite{hollandIntroducingRole2014}, \texttt{prov:specializationOf} \autocite{w3-prov-o}, OAI-ORE Proxy \autocite{ORESpecificationAbstract}) \\
RDA-R1-01M
  & Plurality of accurate and relevant attributes are provided to allow reuse
  & FDOF4
  & Required. Kernel metadata attributes desired \autocite{fdo-KernelAttributes} but not assigned PIDs yet.
  & Unspecified. Multiple metadata records can allow multiple semantic profiles.
  & Large number of general and domain-specific vocabularies can make it hard to find relevant attributes. Rough consensus on kernel metadata: schema.org \autocite{10pzKCEUE}, Dublin Core Terms \autocite{DCMIMetadataTerms}, DCAT \autocite{w3-vocab-dcat-2}, FOAF \autocite{FOAFVocabularySpecification} \\
RDA-R1.1-01M
  & Metadata includes information about the licence under which the data can be reused
  & ---
  & \texttt{licenseConditions} URL/PID in kernel metadata \autocite{fdo-KernelAttributes}
  & ---
  & Dublin Core Terms \texttt{dct:license} frequently recommended, frequently not required, e.g.~\href{https://www.w3.org/TR/vocab-dcat-2/\#Property:distribution_license}{by DCAT 2} \autocite{w3-vocab-dcat-2} \\
RDA-R1.1-02M
  & Metadata refers to a standard reuse licence
  & ---
  & ---
  & ---
  & \href{https://spdx.org/licenses/}{SPDX} and \href{https://creativecommons.org/}{Creative Commons} URIs common, identifiers often inconsistent \\
RDA-R1.1-03M
  & Metadata refers to a machine-understandable reuse licence
  & ---
  & ---
  & ---
  & \href{https://spdx.dev/resources/use/\#documents}{SPDX documents} uncommon \\
RDA-R1.2-01M
  & Metadata includes provenance information according to community-specific standards
  & FDOR9 FDOR10
  & Unspecified (some CORDRA types add getProvenance methods). PID Kernel attributes? Unspecified W3C PROV-O, PAV
  & & \\
RDA-R1.2-02M
  & Metadata includes provenance information according to a cross-community language
  & FDOR9 FDOR8
  & ---
  & ---
  & W3C PROV-O \autocite{w3-prov-o}, PAV \autocite{ciccaresePAVOntologyProvenance2013e}, Dublin Core Terms \autocite{DCMIMetadataTerms} \\
RDA-R1.3-01M
  & Metadata complies with a community standard
  & FDOR10 FROR8
  & (Emerging, e.g.~DiSSCo Digital Specimen \autocite{hardistySpecimenDataRefinery2022a})
  & ---
  & Common, e.g.~DCAT 2 \autocite{w3-vocab-dcat-2}, BioSchemas \autocite{Bioschemas} \\
RDA-R1.3-01D
  & Data complies with a community standard
  & (FDOR3)
  & ---
  & ---
  & Common, HTTP use registered IANA \href{https://www.iana.org/assignments/media-types/media-types.xhtml}{media types}, additional scientific file formats frequently not standardised or identified \\
RDA-R1.3-02M
  & Metadata is expressed in compliance with a machine-understandable community standard
  & FDOF4 FDOF10
  & Recommended
  & ---
  & Common practice for ontologies, specially in bioinformatics, e.g.~BioPortal \autocite{NCBOBioPortal}, Darwin Core \autocite{wieczorekDarwinCoreEvolving2012} \\
RDA-R1.3-02D
  & Data is expressed in compliance with a machine-understandable community standard
  & (FDOR2)
  & No, FDO is typed but data can be any bytestream
  & ---
  & Occassionally, (e.g.~\href{https://github.com/The-Sequence-Ontology/Specifications/blob/master/gff3.md}{GFF3}, \href{https://fits.gsfc.nasa.gov/fits_standard.html}{FITS}, \href{https://www.loc.gov/preservation/digital/formats/fdd/fdd000280.shtml}{ESRI}) \\
\bottomrule
\end{longtable}
\end{landscape}