### 2023-05-18 latex/table6.tex
simplified table latex

\begin{landscape}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.0833}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3056}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3056}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3056}}@{}}
\caption[Assessing EOSC Interoperability Framework, against FDO \& Linked Data]{Assessing EOSC Interoperability Framework \autocite[section 3.6]{corchoEOSCInteroperabilityFramework2021b} against the FDO guidelines \autocite{RwvirqWg} and Linked Data practices.}
\label{tbl:eosc}\tabularnewline
\toprule
Layer &
Recommendation &
FDO &
Linked Data \\
\midrule
\endhead
Technical      & Open Specification 
  & FDO specifications are semi-open, process gradually more transparent 
  & Open and transparent standard processes through W3C \& IETF \\
Technical      & Common security \& privacy framework 
  & Unspecified 
  & TLS for encryption, multiple approaches for single-sign-on (e.g.~ORCID, Life Science Login). Privacy largely unspecified. \\
Technical      & Easy SLAs for service providers 
  & Unspecified 
  & None \\
Technical      & Access data in different formats 
  & None formalised, custom operations or relations 
  & Content-negotiation, \texttt{rel=alternate} relations \\
Technical      & Coarse-grained/fine-grained search tools 
  & Freetext \texttt{0.DOIP/Op.Search} on local DOIP, no federation 
  & Coarse-grained e.g.~\href{https://datasetsearch.research.google.com/}{Google Dataset Search}, fine-grained (e.g.~federated SPARQL) require detailed vocabulary/metadata insight \\
Technical      & Clear PID policy 
  & Strong FDO requirements, tends towards Handle system. 
  & Not required, different communities set policies \\
Semantic       & Clear definitions for concepts/metadata/schemas 
  & Required by FDO requirements, but not yet formalised 
  & Ontologies, SKOS, OWL \\
Semantic       & Semantic artefacts w/ open licenses 
  & All artefacts are PIDs, license not yet required by kernel metadata
  & Open License is best practice for ontology publishing \\
Semantic       & Documentation for each semantic artefact 
  & No direct rendering from FDO, no requirement for human-readable description 
  & Ontology rendering, content-negotiation \\
Semantic       & Repositories of artefacts 
  & Required, but not formalised 
  & Bioontologies, otherwise not usually federated \\
Semantic       & Repositories w/ clear governance 
  & Recommended 
  & Largely self-governed repositories, if well-established may have clear governance. \\
Semantic       & Minimal metadata model for federated discovery 
  & Kernel metadata \autocite{fdo-KernelAttributes} based on RDA recommendations \autocite{weigelRDARecommendationPID2018}.
  & DCAT, schema.org, Dublin Core \\
Semantic       & Crosswalks from minimal metadata model 
  & FDO Typing recommends referencing existing type definitions, but not as separate crosswalks 
  & Multiple crosswalks for common metadata models, but frequently not in semantic format \\
Semantic       & Extensibility options for diciplinary metadata 
  & Communities encouraged to establish own types 
  & Extensible by design, domain-specific metadata may be at different granularity \\
Semantic       & Clear protocols/building blocks for federation/harvesting of artefact catalogues 
  & Collection types not yet defined 
  & SWORD, OAI-PMH \\
Organisational  & Interoperability-focused rules of participation recommendations 
  & Recommended 
  & Implied only by some communities, tendency to specialise \\
Organisational  & Usage recommendations of standardised data formats 
  & None 
  & None -- but common for metadata (e.g.~JSON-LD) \\
Organisational  & Usage recommendations of vocabularies 
  & Recommended by community 
  & Common (see \href{https://rdmkit.elixir-europe.org/metadata_management}{RDMKit}) \\
Organisational  & Usage recommendations of metadata 
  & Recommended by community 
  & RO-Crate, Bioschemas \\
Organisational  & Management of permanent organization names/functions 
  & Handle owner, but unclear contact. Contact info in DOIP service provider 
  & ROR. DCAT contacts. \\
Legal          & Standardised human and machine-readable licenses 
  & None 
  & \href{https://spdx.org/licenses/}{SPDX}, but not that frequently used \\
Legal          & Permissive licenses for metadata (CC0, CC-BY-4.0) 
  & Undefined 
  & Both CC0, CC-BY-4.0 common, e.g.~in DCAT \\
Legal          & Different licenses for different parts 
  & Each part as separate FDO can have separate license 
  & DCAT, RO-Crate, Named graphs for splitting metadata \\
Legal          & Mark expired/inexistent copyright 
  & Undefined 
  & Unclear, semantics assume copyright valid \\
Legal          & Mark orphaned data 
  & Tombstone for deleted data, but no owner of DOIP server means FDO disappears 
  & Frequently data and endpoint has no known maintainer, archiving in common repositories becoming common \\
Legal          & List recommended licenses 
  & Undefined 
  & Best practice recommendations \\
Legal          & Track license evolution for dataset 
  & Undefined 
  & Versioning with PAV/PROV/DCAT \\
Legal          & Policy/guidance for patent/trade secrets violation 
  & Undefined 
  & Undefined, legal owner may be specified \\
Legal          & GDPR compliance for personal data 
  & Undefined 
  & Undefined \\
Legal          & Restrict access/use if legally required 
  & By transport protocol (undefined by FDO/DOIP) 
  & Diverging approaches, typically landing pages w/ auth\&auth or click-thru \\
Legal          & Harmonised terms-of-use 
  & Undefined 
  & Undefined \\
Legal          & Alignment between EOSC and national legislation 
  & Not applicable 
  & Not applicable \\
\bottomrule
\end{longtable}
\end{landscape}