### 2023-06-12 latex/table2.tex
40 pages!

\begin{table}[h!]
\centering
\caption{Mapping the Metamodel concepts from the Interoperability Framework for Fast Data \autocite{delgadoInteroperabilityFrameworkDistributed2016a} to equivalent concepts for FDO and Web.
\label{tbl:metamodel-concepts}}\tabularnewline
 \begin{tabular}{ m{5em}  m{15em} m{15em} } 
 \hline
Metamodel concept & 
FDO/DOIP concept & 
Web/HTTP concept \\ 
 \hline
Resource	  & FDO/DO	            & Resource \\
Service	    & DOIP service	      & Server/endpoint \\
Transaction	& (not supported)	    & Conditional requests, \texttt{409\ Conflict} \\
Process	    & Extended operations	& (primarily stateless), \texttt{100\ Continue}, \texttt{202\ Accepted} \\
Operation	  & DOIP Operation	    & Method, query parameters \\
Request	    & DOIP Request	      & Request \\
Response	  & DOIP Response	      & Response \\
Message	    & Segment, \texttt{requestId} 
                                  & Message, Representation \\
Channel	    & DOIP Transport protocol (e.g.~TCP/IP, TLS). JSWS signatures.
                                  & TCP/IP, TLS, UDP \\
Protocol  	& DOIP 2.0, ++	      & HTTP/1.1, HTTP/2, HTTP/3 \\
Link	      & PID/Handle	        & URL \\ 
 \hline
 \end{tabular}\end{table}